% Maybe talk about scope instead of purpose?
The main purpose of the engine is to render 3D scenes.
Nevertheless, it should be possible to use the tools to render 2D scenes as well,
though using the engine in this way is left as an exercise to the reader.
In order to provide comprehensive tools for developers,
the scope of the project encompasses the following aspects of 3D rendering:

\begin{itemize}
  \item \textbf{Bitmap Font Rendering.} A very helpful module that will allow a user to render debug information directly on the screen in real time.
  \item \textbf{3D Model Rendering.} The most essential feature for rendering 3D scenes. The engine is able to import 3D meshes from Wavefront OBJ files and render them.
  \item \textbf{Texture Mapping.} Furthermore, users can import PNG texture files, and map them to meshes for further detail and aesthetics.
  \item \textbf{Lighting.} In order to provide a better depth perception, the Phong reflection model is implemented to model different types of lights.
  \item \textbf{Procedural Generation.} While not a fundamental aspect of 3D rendering, different algorithms like Perlin Noise and Marching Cubes are implemented to allow a user to render arbitrarily large or detailed scenes.
\end{itemize}

Note that the scope of the project is far from exhaustive.
There are many other aspectes of graphics rendering like shadow mapping, physics based rendering, physics simulations and many others\cite{learnopengl}.
While they are outside of the scope of this project,
the hope is that the implementation will allow these features to be integrated in future iterations without much additional effort.
